---
title: "p8105_hw5_jw4348"
author: "Jingyu Wang"
output: github_document
date: "2023-11-15"
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

For this problem, we are interested in data gathered and made public by _The Washington Post_ on homicides in 50 large U.S. cities. The code chunk below imports and cleans the data.

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )
  ) %>% 
  filter(city_state != "Tulsa, AL") 
```

The resulting dataframe has `r nrow(homicide_df)` entries, on variables that include the victim name, race, age, and sex; the date the homicide was reported; and the location of the homicide. In cleaning, I created a `city_state` variable that includes both city and state, and a `resolution` variable to indicate whether the case was closed by arrest. I also excluded one entry in Tulsa, AL, which is not a major US city and is most likely a data entry error. 

In the next code chunk, I group within cities and summarize to produce the total number of homicides and the number that are solved. 

```{r}
city_homicide_df = 
  homicide_df %>% 
  select(city_state, disposition, resolution) %>% 
  group_by(city_state) %>% 
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolution == "unsolved"))
```

Focusing only on Baltimore, MD, I can use the `prop.test` and `broom::tidy` functions to obtain an estimate and CI of the proportion of unsolved homicides in that city. The table below shows those values.

```{r}
bmore_test = 
  prop.test(
    x = filter(city_homicide_df, city_state == "Baltimore, MD") %>% pull(hom_unsolved),
    n = filter(city_homicide_df, city_state == "Baltimore, MD") %>% pull(hom_total)) 

broom::tidy(bmore_test) %>% 
  knitr::kable(digits = 3)
```

Building on this code, I can use functions in the `purrr` package to obtain estimates and CIs for the proportion of unsolved homicides in each city in my dataset. The code below implements this analysis. 

```{r}
test_results = 
  city_homicide_df %>% 
  mutate(
    prop_tests = map2(hom_unsolved, hom_total, \(x, y) prop.test(x = x, n = y)),
    tidy_tests = map(prop_tests, broom::tidy)) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high) %>% 
  mutate(city_state = fct_reorder(city_state, estimate))
```

Finally, I make a plot showing the estimate (and CI) of the proportion of unsolved homicides in each city.

```{r}
test_results %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This figure suggests a very wide range in the rate at which homicides are solved -- Chicago is noticeably high and, given the narrowness of the CI, likely is the location of many homicides. 

## Problem 2

I will create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time.

#### Start with a dataframe containing all file name.Iterate over file names and read in data for each subject and saving the result as a new variable in the dataframe
```{r}
df_study = 
  tibble(
    file = list.files("data/problem_1/"),
    path = str_c("data/problem_1/", file)
        ) |> 
    mutate(data = map(path, read_csv)) |>
    unnest()
df_study
```

#### And then I tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
tidy_df = 
  df_study |> 
  mutate(
    file = str_remove(file, ".csv"),
        ) |>
  separate(file, into = c("arm", "subject_id"), sep = "_") |>
  pivot_longer(
    cols = week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observations"
              ) |>
  select(subject_id, arm, week, observations)
tidy_df
```

#### Next I will make a spaghetti plot showing observations on each subject over time

```{r}
tidy_df |> 
  ggplot(aes(x = week, y = observations, group = subject_id, color = arm)) +
  geom_point() +
  geom_path() +
  facet_grid( ~ arm) +
  labs(title = "Observations on Each Subject over Time",
       x = "Week",
       y = "Observations") 
```

#### Finally, I will comment on differences between groups.
- In the control arm:
  - The observations fluctuate without a clear overall trend.
  There is no consistent upward or downward pattern in observations from week 1 to 8.

- In the experiment arm:
  - The overall trend in observations is upward over time.
  - Unlike the control group, there is a noticeable increase in observations from week 1 to 8.

- Therefore, the difference trends between groups suggests a potential difference in the response to treatment between the control and experiment groups, with the experiment group showing a more systematic increase in observations over time.